# Assignment 1: Reproducible Research

*Fazli Haleem*

*December 23, 2018

# Introduction

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals throughout the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

Loading required packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(lubridate)
library(ggplot2)
```

## setting working directory and loading data:

```{r cars}
setwd("//id.aau.dk/Users/haleem/Desktop/Assignment1")
data <- read.csv("activity.csv")
head(data)
str(data)
# tidying date
data$date <- ymd(data$date)
(echo = TRUE)
```

# What is mean total number of steps taken per day?

## 1)CALCULATING THE STEPS TAKEN

```{r}
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

## 2)CONSTRUCTING THE PLOT USING ggplot

```{r}
ggplot(steps, aes(x = steps)) +
  geom_histogram(fill = "red", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

# 3)CALCULATING THE MEAN AND THE MEDIAN

```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
```

# What is the average daily activity pattern?

## CALCULATING THE AVERAGE STEPS

```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
```

## constructing graph

```{r}
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "firebrick")

```
# FINDING THE MAX STEPS:
```{r}
interval[which.max(interval$steps),]
```

# Imputing missing values

## Filling missing values

```{r}
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```

#Calculatin the steps

```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

# PLOTTING THE GRAPH

```{r}
ggplot(steps_full, aes(x = steps)) +
  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```

# CALCULATING THE MEAN AND THE MEDIAN

```{r}
mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
```

# Are there differences in activity patterns between weekdays and weekends?

## WHETHER WEEKDAY OR WEEKEND:

```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```

# AVERAGE STEPS IN 5 MINS INTERVALS:

```{r}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
```

# PLOTTING THE GRAPH:

```{r}
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
